---
title: "DoD Military Community and Family Policy Military OneSource eNewsletter"
author: "Paul Testa and Jake Bowers"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
    html_document:
        toc: TRUE
	template: cosmo
---

```{r init,echo=F}
library(knitr)
opts_chunk$set(eval=T,echo=F,results="hide",message=F,warning=F,cache=T)

```

\tableofcontents

# Setup

- Download "designdata.csv" and "subscriptiondata.csv" onto local machine
- Load files and libaries

```{r setup}
wrkdat<-read.csv("data/designdata.csv",as.is=TRUE)
subscribers<-read.csv("data/subscriptiondata.csv",as.is=TRUE)
```


# Data

First check data

```{r clean}
stopifnot(nrow(wrkdat)==491879) # 491879 observations

# Make sure emails are unique:
stopifnot(length(unique(wrkdat$emailhash))==dim(data)[1])
stopifnot(length(unique(subscribers$emailhash))==dim(subscribers)[1])

stopifnot(nrow(subscribers)==7759) # 7759 subscriptions
stopifnot(sum(wrkdat$emailhash%in%subscribers$emailhash)==5563) # 5563 subscribers in initial email list

# Rename Treatment indicator
names(wrkdat)[3]<-"treatment"
```

# Mark unmatchable subscribers

If a subscriber has no match on email or full name or last name to any email or name in the experimental pool, note this. These people are unmatchable. They may be people who subscribed but were outside the study. Or they may be people in the study who subscribed changing their names and email addresses. For now, we exclude these people from matching. We'll return to them later.

```{r}
subscribers$noemail2match <- !(subscribers$emailhash %in% wrkdat$emailhash)
subscribers$noname2match <-  !(subscribers$name1hash %in% na.omit(unlist(wrkdat[,grep("name[0-9].?hash",names(wrkdat),value=TRUE)])))
subscribers$nolastname2match <-  !(subscribers$Last.Name1hash %in% na.omit(unlist(wrkdat[,grep("Last.Name[0-9].?hash",names(wrkdat),value=TRUE)])))
subscribers$noway2match <- subscribers$noemail2match & subscribers$noname2match & subscribers$nolastname2match
stopifnot(sum(subscribers$noway2match)==177)
```

There are `r sum(subscribers$noway2match)` people who have no email, no name,
and no last name that overlap with anyone in the original experimental pool
(this includes looking accross up to 24 different names and up to 24 different
last names). If these people subscribed but were not part of the experiment, we
can just exclude them. If these people were part of the experiment, then we
have 744 positive outcomes with unknown relationship with treatment assignment.
The question we will ask is whether our main results will change (and how much)
given a few ways to distribute these responses across treatments and
batches.

## Match on email

Then,  match those sent email treatments to subscribers by email address.

```{r}
# Matches by email
stopifnot(sum(wrkdat$emailhash%in%subscribers$emailhash)==5563) # 5563

# Matches by name1hash
stopifnot(sum(wrkdat$name1hash%in%subscribers$name1hash)==7951) # 7951 subscribers in

# Additional variables
## table(wrkdat$Order.Within.Condition)
table(wrkdat$Condition.Order.Within.Batch)

### Make sure that each row is a unique email address
stopifnot(nrow(wrkdat)==length(unique(wrkdat$emailhash)))
stopifnot(nrow(subscribers)==length(unique(subscribers$emailhash)))
row.names(wrkdat)<-wrkdat$emailhash
row.names(subscribers)<-subscribers$emailhash

## Make sure that Subscribe.Date has no missing values or extreme values.
stopifnot(all(!is.na(subscribers$Subscribe.Date)))
sort(unique(subscribers$Subscribe.Date))[1:10]
rev(sort(unique(subscribers$Subscribe.Date)))[1:10]
```

First match on email addresses. If an email address in the design data ends up
in the subscriber file, then we know that person subscribed to the newsletter.

```{r}
matchedemails<-intersect(row.names(wrkdat),row.names(subscribers))
wrkdat[matchedemails,"sdate"]<-subscribers[matchedemails,"Subscribe.Date"]
stopifnot(sum(!is.na(wrkdat$sdat))==5563)

subscribers$noemailmatch<- !(row.names(subscribers) %in% matchedemails)
stopifnot(sum(!subscribers$noemailmatch)==5563)

## Excluding the people for whom there is no point in matching for now
subdat<-subscribers[subscribers$noemailmatch & !subscribers$noway2match,]

```

So, we have 2019 subscribers who were not matched on email address.

## Match on full name within batch

Next, among those 2169 unmatched subscribers, try to match on name within
batch. This is tricky but we start simple.  Add batch information to the
subscription file to add in merging otherwise we will have *at least* 900
subscriptions with no match to the experimental treatments.

We assume that if you have subscribed after the mailing of batch A but before
batch B, then you were randomized to batch A.

```{r}
table(wrkdat$Batch)
table(as.Date(subscribers$Subscribe.Date,format="%m/%d/%y"),exclude=c())
```

Add start date of batch using information provided by the agency.^[After cleaning the file in a spreadsheet to remove extraneous information.]

```{r}
batchdates<-read.csv("batchdates.csv",as.is=TRUE,header=TRUE,strip.white=TRUE)
## Remove empty rows
batchdates[batchdates==""]<-NA ## replace empty strings with NA
emptyrow<-apply(batchdates,1,function(x){ all(is.na(x))})
badrows<-apply(batchdates,1,function(x){ tmp<-grep("TAR",x); if(length(tmp)==0){ return(FALSE) } else { TRUE } })
batchdates<-batchdates[!emptyrow & !badrows,]
batchdates$Batch<-rep(1:82,each=2)
stopifnot(all(na.omit(batchdates$Batch.Number-batchdates$Batch)==0))

## Fill in treatment information
batchdates$eMail.to.send[seq(2,164,by=2)]<-batchdates$eMail.to.send[seq(1,163,by=2)]
batchdates$eMail.to.send.1[seq(2,164,by=2)]<-batchdates$eMail.to.send.1[seq(1,163,by=2)]
batchdates$eMail.to.send.2[seq(2,164,by=2)]<-batchdates$eMail.to.send.2[seq(1,163,by=2)]
batchdates$eMail.to.send.3[seq(2,164,by=2)]<-batchdates$eMail.to.send.3[seq(1,163,by=2)]
batchdates$eMail.to.send.4[seq(2,164,by=2)]<-batchdates$eMail.to.send.4[seq(1,163,by=2)]
batchdates$eMail.to.send.5[seq(2,164,by=2)]<-batchdates$eMail.to.send.5[seq(1,163,by=2)]

## Pull off the actual send times and change the data to have multiple rows for batches rather than columns
bdlong<-reshape(batchdates,direction="long",v.names=c("eMail.to.send","Details"),
		varying=list(eMail.to.send=grep("eMail.to.send",names(batchdates),value=TRUE),
			     Details=grep("Details",names(batchdates),value=TRUE)))

bdlongActual <- bdlong[grep("Actual",bdlong$Details),]

## Three actual send times were not recorded (we assume they were actually sent). So use the set time.
##bdlongActual[bdlongActual$Batch==16&bdlongActual$eMail.to.send=="A",] <- bdlong[bdlong$Batch==16&bdlong$eMail.to.send=="A"&bdlong$Batch.Number==16,"Details"]
## Strip off "PM"
bdlongActual[bdlongActual$Batch==16&bdlongActual$eMail.to.send=="A","Details"] <- bdlong["31.4","Details"]
bdlongActual[bdlongActual$Batch==36&bdlongActual$eMail.to.send=="E","Details"] <- bdlong["71.5","Details"]
bdlongActual[bdlongActual$Batch==80&bdlongActual$eMail.to.send=="F","Details"] <- bdlong["159.6","Details"]

## Make sure that each treatment is associated with one and only one batch
stopifnot(all(with(bdlongActual,table(Batch,eMail.to.send))==1))

## Convert Details into dates
bdlongActual$dateString<-sapply(strsplit(bdlongActual$Details,":\ "),function(x){ x[2] })
bdlongActual$date <- as.Date(bdlongActual$dateString,format="%d-%b-%Y %H:%M:%S")
bdlongActual[grep("PM",bdlongActual$dateString),"date"]<-as.Date(bdlongActual$dateString[grep("PM",bdlongActual$dateString)],
								 ,format="%m/%d/%Y %H:%M:%S %p")
stopifnot(sum(is.na(bdlongActual$date))==0)
table(bdlongActual$date)
```

Now, the subscription data only has day of subscription, no time. So, in order
to map date of subscription onto batch among those not matched by email, we
have to collapse batches into a day (often there were two batches per day). We
hope that we can still back out original batch after doing the matching by name
and collapsed-batch. If not, we will lose some statistical power because our
blocking variable (batch) will be less fine grained, but we will probably gain
more because we can increase the sample size and also will reduce our problems
with unmatchable subscribers (unknown relationships between treatement
assignment and outcome).

```{r}
subdat$dateDay<-as.Date(subdat$Subscribe.Date,format="%m/%d/%y")

## Batch Day/Collapsed Batch to Date mapping
bdlongActual$dateDay <- round(bdlongActual$date)
stopifnot(length(unique(bdlongActual$dateDay))==41)

batchdaydat<-data.frame(dateDay=unique(bdlongActual$dateDay),
			batchday=rank(unique(bdlongActual$dateDay)))
row.names(batchdaydat)<-batchdaydat$dateDay

## Just add appropriate batchday info to the subdat data
subdat<-merge(subdat,batchdaydat,all.x=TRUE,sort=FALSE)
## Now we have some subscription dates that are not on a mailing day
sum(is.na(subdat$batchday))
stopifnot(length(unique(subdat$emailhash))==nrow(subdat))
row.names(subdat)<-subdat$emailhash

intermaildates<-unique(subdat$dateDay[is.na(subdat$batchday)])
## Choose the mailing date that is closest to the subscription day, yet before the subscription day

getcloseday<-function(x){
	# x is date subscribed, a scalar
	y <- batchdaydat$dateDay
	thediff<-x - y
	return( as.Date(max(y[thediff>=0])) )
}

getcloseday(subdat$dateDay[1])

## This is a very weird hack to ensure that the newdate column stayed a Date
newdate <- vector(length=nrow(subdat))
class(newdate)<-"Date"
for(i in 1:length(subdat$dateDay)){
	newdate[i]<-getcloseday(subdat$dateDay[i])
}
subdat$newdate<-newdate

stopifnot(all(with(subdat[!is.na(subdat$batchday),],dateDay-newdate)==0))

subdat$batchday2<-NA
subdat$batchday2 <- batchdaydat[as.character(subdat$newdate),"batchday"]

## Check that each newdate is associated with one batch and only one batch
thetab<-with(subdat,table(newdate,batchday2))
stopifnot(all(c(unique(thetab[upper.tri(thetab)]), unique(thetab[lower.tri(thetab)]))==0))
stopifnot(nrow(subdat)==2019)
```

Now, make a collapsed batchday variable on the design data:

```{r}
wrkdat$batchday2<-as.numeric(cut(wrkdat$Batch,seq(0,82,by=2)))
## Make sure that the new batch collects exactly two of the original batches
tmptab<-with(wrkdat,table(batchday2,Batch))
stopifnot(all(apply(tmptab,1,function(x){ sum(x!=0) }) == 2 ))
stopifnot(all(apply(tmptab,2,function(x){ sum(x!=0) }) == 1 ))
```

Now match subscribers who have the same name *and* batch number as experimental pool subjects.

```{r}
## Make a dataset called expool for those people in the experimental pool not matched on email
expool<-wrkdat[is.na(wrkdat$sdate),]
stopifnot(nrow(expool)==486316)

## Go batch by batch just for clarity. There are probably faster vectorized ways.
lastnms<-grep("Last.Name[0-9].?hash$",names(wrkdat),value=TRUE)
expbatches <- split(expool[,c("emailhash","name1hash","sdate","batchday2",lastnms)],expool$batchday2)
subbatches <- split(subdat,subdat$batchday2)
stopifnot(length(expbatches)==length(subbatches))

## Remove subscriber names that are duplicated within batch
uniqsubs<-lapply(subbatches,function(dat){
			 thetab <- table(dat$name1hash)
			 dups <- names(thetab[thetab>1])
			 if(length(dups)==0){
				 message("ok")
				 return(dat)
			 } else {
				 message(length(dups))
				 return(dat[!(dat$name1hash %in% dups),])
			 }
			})
## How many names do we remove? About 4*2=8.
table(sapply(subbatches,nrow)-sapply(uniqsubs,nrow))

## Now merge from the subscribers onto the experimental pool batch by batch
merge2list<-list()
for(i in 1:length(expbatches)){
	message(i)
	merge2list[[i]]<- merge(expbatches[[i]],uniqsubs[[i]],by="name1hash",all.x=TRUE,sort=FALSE)
	stopifnot(nrow(merge2list[[i]])==nrow(expbatches[[i]]))
	uniqsubs[[i]]$matched <- uniqsubs[[i]]$name1hash %in% expbatches[[i]]$name1hash
}
sum(sapply(merge2list,nrow))
mergeexpool<-do.call("rbind",merge2list)
mergeexpool$sdate<-mergeexpool$Subscribe.Date
subdat2<-do.call("rbind",uniqsubs)
subdat2<-subdat2[!subdat2$matched,]
stopifnot(nrow(subdat2)==1206)
rm(expbatches,uniqsubs,merge2list,i,subbatches)
```

So we have 1206 subscribers who are not matched after matching on full name and email address.

## Match on last name within batch

Now, do one last matching on unique last names within batch. Some inspection of
the data suggested that people called "Jake Bowers" might sometimes use "J
Bowers" or "Jacob Bowers". So, among those not already matched, and among those
last names that are unique within batch, do one last merge.

```{r}
## First remove last names that are duplicated within batch among subscribers (about 39)
subdat2$duplastname<-unsplit(lapply(split(subdat2,subdat2$batchday2),function(dat){
				    thetab <- table(dat$Last.Name1hash)
				    dups <- names(thetab[thetab>1])
				    dat$Last.Name1hash %in% dups
			     }),subdat2$batchday2)
sub2batches<-split(subdat2[!subdat2$duplastname,],subdat2[!subdat2$duplastname,"batchday2"])

mergeexpool$Last.Name1hash<-mergeexpool$Last.Name1hash.x
mergeexpool$emailhash<-mergeexpool$emailhash.x
mergeexpool$batchday2<-mergeexpool$batchday2.x
expool2batches<-split(mergeexpool[,c("emailhash",lastnms,"batchday2","sdate")],mergeexpool$batchday2)

## Find the subscriber last names which uniquely match *any* of the experimental pool last names within batchday
whichsub2list<-list()
for(i in 1:length(expool2batches)){
	message(i)
	whichsub2list[[i]]<-apply(expool2batches[[i]][,lastnms],1,function(x){
					  ##res<-sub2batches[[i]]$Subscribe.Date[sub2batches[[i]]$Last.Name1hash %in% x]
					  res<- which(sub2batches[[i]]$Last.Name1hash %in% x)
					  if(length(res)==0 | length(res)>1){
						  ## Not a match if no match or if more than one unique match
						  return(NA)
					  } else {
						  return(res)
					  }
			})
        thetab <- table(whichsub2list[[i]])
	uniqnms <- as.numeric(names(thetab[thetab==1]))
	poolrows <- names(whichsub2list[[i]][whichsub2list[[i]] %in% uniqnms])
	expool2batches[[i]]$sdate3<-NA
	expool2batches[[i]][poolrows,"sdate3"]<- sub2batches[[i]][uniqnms,"Subscribe.Date"]
	sub2batches[[i]]$nolastnmmatch<-TRUE
	sub2batches[[i]][uniqnms,"nolastnmmatch"]<-FALSE
}

subdat3<-do.call("rbind",sub2batches)
stopifnot(nrow(subdat2[!subdat2$duplastname,])==nrow(subdat3))

mergeexpool2 <- do.call("rbind",expool2batches)
stopifnot(nrow(mergeexpool2)==nrow(mergeexpool))

rm(expool2batches,sub2batches)
mergeexpool2$finalsdate<-with(mergeexpool2,ifelse(is.na(sdate3),sdate,sdate3))
table(is.na(mergeexpool2$finalsdate))
table(is.na(mergeexpool2$sdate))
table(is.na(mergeexpool2$sdate3))

stopifnot(length(unique(mergeexpool2$emailhash))==nrow(mergeexpool2))
row.names(mergeexpool2)<-mergeexpool2$emailhash.x

```

Which subscribers are still not matched? What (pseudo)batches were they in?

```{r}


```

## Put it all back together

Put the experimental pool back together.

```{r}

## Make sure same number of rows.
```
