---
title:Data Setup and Cleaning
author: Jake Bowers
date: 7 July 2016
...

The construction of the outcome measure depends on matching email addresses of those who subscribed with those who were randomly assigned to different conditions.

 - Downloaded the `01 mos_email_addresses.20150508122202 - randomized to MC&FP 18 NOV (2).xlsx` and `SBST_subscriptions_03182016.docx.xlsx` files from the emails to which they had been attached by the agency contacts and decrypted it using the provided password.
 - Converted those files to .csv by hand using Microsoft Excel

First, combine the two files containing outcome information:

```{r}
subscriptions0<-read.csv("SBST_subscriptions_03182016.sheet0.csv",as.is=TRUE,header=TRUE)
subscriptions1<-read.csv("SBST_subscriptions_03182016.sheet1.csv",as.is=TRUE,header=TRUE)
subscriptions <- rbind(subscriptions0,subscriptions1)
## We should have only one unique email address per row
stopifnot(length(unique(subscriptions$Email.Address))==nrow(subscriptions))
```

Second, import the design information:
```{r}
rand <- read.csv("01 mos_email_addresses.20150508122202 - randomized to MC&FP 18 NOV (2).csv",as.is=TRUE,header=TRUE)
stopifnot(length(unique(rand$Email.Address))==nrow(rand))

```

Third, generate hashes. The idea is to create a unique identifier for each person. Since email addresses were used originally for this purpose, and we have shown that they are, in fact, unique, we'll try to convert them into some format that retains their uniqueness but from which the emails themselves cannot (basically ever) be recovered.

For example, here is how hashes work:
```{r}
library(openssl)
emailaddresses <- c("jacob.bowers@gsa.gov","jwbowers@illinois.edu")
hashedaddresses <- sha256(emailaddresses)
cbind(hashedaddresses,emailaddresses)
```

Now, do this for the full data. Create key files mapping emails to hashes just in case we have made a mistake or want to double check something. These files will be stored on the secure Google Drive along with the original files. Then create anonymized files that we can use to create the final analysis file.

```{r}
subscriptions$id <- sha256(subscriptions$Email.Address)
rand$id <- sha256(rand$Email.Address)

randkey <- rand[,c("Email.Address","id")]
designdata <- rand[,which(names(rand) != "Email.Address")]

str(randkey)
str(designdata)

subscriptionkey <- subscriptions[,c("Email.Address","id")]
subscriptiondata <- subscriptions[,c("id","Subscribe.Date")]

str(subscriptionkey)
str(subscriptiondata)

write.csv(subscriptionkey,file="subscriptionkey.csv")
write.csv(randkey,file="randkey.csv")
write.csv(subscriptiondata,file="subscriptiondata.csv")
write.csv(designdata,file="designdata.csv")
```
