
---
title: "DoD Military Community and Family Policy Military OneSource eNewsletter: Google Analytics Results"
author: "Paul Testa and Jake Bowers"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
    html_document:
      toc: TRUE
---

```{r init,echo=F, results=F}
## Easy way to look for and install missing packages and load them
if (!require("pacman")){ install.packages("pacman") }
pacman::p_load("knitr","openssl","mosaic","dplyr","coin","ggplot2","sandwich","lmtest","reshape2","multcomp","lubridate")

opts_chunk$set(tidy=TRUE,echo=TRUE,results='markup',strip.white=TRUE,cache=FALSE,highlight=TRUE,width.cutoff=132,size='footnotesize',message=FALSE,warning=TRUE,comment=NA)

options(digits=4,width=100,scipen=8)
```

\tableofcontents

# Setup

This chunk records the series of steps we took to clean and collapse the Google Analytics data after converting it to csv via excel.

```{bash, eval=FALSE}
## Assuming the bash or zsh shell. This is not R code
cp SBST-Batch-SendTracking-GAStats-15Aug2016.csv orig.csv
sed -i "" "s/\*no //g" SBST-Batch-SendTracking-GAStats-15Aug2016.csv
## By Hand: Added a header
sed -i "" 's/\/15,.*$/\/15,/' SBST-Batch-SendTracking-GAStats-15Aug2016.csv
sed -i "" 's/\/16,.*$/\/16,/' SBST-Batch-SendTracking-GAStats-15Aug2016.csv
gsed -i.bak '/^[0-9]/! s/^/,/' SBST-Batch-SendTracking-GAStats-15Aug2016.csv ## add extra column, use gnu-sed from brew install gnu-sed
gsed -i.bak 's/reported/\/ email/g' SBST-Batch-SendTracking-GAStats-15Aug2016.csv
```

The version of `SBST-Batch-SendTracking-GAStats-15Aug2016.csv` that is on Google Drive was cleaned as above. To execute this file, you'll need to download it.


```{r}
gadat<-read.csv("data/SBST-Batch-SendTracking-GAStats-15Aug2016.csv",as.is=TRUE,header=TRUE)
gadat$Source...Medium[gadat$Source...Medium==""]<-NA
table(gadat$Source...Medium)
```

Google Analytics recorded data for each newsletter treatment for 103 days from the end of Nov 2015 to early March 2016.

```{r}
range(mdy(gadat$X[gadat$X!=""]))
```

This next just reorganizes the data for easier analysis.

```{r}
tabdat <- gadat %>% group_by(Source...Medium) %>% summarise(sum(New.Users,na.rm=TRUE))
names(tabdat) <- c("treatment","newusers")
tabdat <- tabdat[-7,]
wrkdat<-read.csv("data/wrkdat.csv",as.is=TRUE)
assigndat<-wrkdat %>% group_by(treatment) %>% summarize(sent=n())

dat <- cbind(assigndat,newusers=tabdat$newusers)
stopifnot(tabdat[tabdat$treatment=="newsletter_a / email","newusers"] == dat[dat$treatment=="A","newusers"])

dat$success <- dat$newusers
dat$failures <- dat$sent - dat$newusers

mat <- as.matrix(dat[,c("success","failures")])
rownames(mat) <- as.character(dat$treatment)

```

Expand these tables to make analyses easier with existing software:

```{r}
bigdat <- melt(apply(mat,1,function(x){ c(rep(1,x[1]),rep(0,x[2])) }))
names(bigdat) <- c("subscribed","treatment")
bigdat$subscribedF <- factor(bigdat$subscribed,levels=c(1,0))
bigdat$treatmentF <- factor(bigdat$treatment)
tab2<-with(bigdat,table(treatment,subscribedF))
```

# Test for any difference across treatments

Now, assess the hypotheses of no difference among treatments (comparing the
approximate chisq test that does not rely on large sample assumptions to the
one that relies on large sample assumptions). We have a lot of evidence that
the different treatments had different effects.

```{r}
prop.test(x=tab2[,1],n=rowSums(tab2))
chisq_test(subscribedF~treatmentF,data=bigdat) ## or chisq_test(tab2)
## This next just checks to see if our reliance on the central limit theorem is warranted in a situation with such rare outcomes.
chisq_test(subscribedF~treatmentF,data=bigdat,distribution=approximate(B=1000))
```
# Which treatment was most effective?

Treatment F (the "Active") treatment was clearly the most effective at
increasing the number of new devices or web browsers loading the first page of
the newsletter site with roughly 2.3% of those sent this treatment accessing
the website compared to, say, about 1.4% for treatment C. Note that this is
only the same as increasing the number of people responding to the email if
each person only reached the web page once. More likely these treatments
motivated people to both click on the link but also to try to use different
devices to reach the site --- perhaps to show it to others, or if it didn't
load well on their preferred device or over the network they were using when
they read the email.


```{r}
thelm <- lm(subscribed~treatmentF,data=bigdat)
c(coef(thelm)[1], coef(thelm)[1]+coef(thelm)[2:6])
coeftest(thelm,vcov=vcovHC(thelm,type="HC2"))
allcomps <- glht(thelm,linfct=mcp(treatmentF="Tukey"))
sumallcomps<- summary(allcomps)
allcompsci <- confint(allcomps) ## adjusted for multiple comparisons
```

We see from the next figure that `r sum(sumallcomps$test$pvalues<0.05)` out of the `r choose(6,2)` pairwise comparisons are statistically distinguishable from 0 at $\alpha=0.05$. Specifically, the effects for Treatment F are significantly larger than any other treatment except Treatment B --- and there the 95% confidence interval just barely includes 0. The effect for Treatment B, although larger than the remaining Treatments (A,C,D and E) is only distinguishable from Treatment C.

```{r figTukey, cache=FALSE, results=TRUE}
# Get estimates and intervals
tukey<-data.frame(allcompsci$confint)
##plot(TukeyHSD(a1))
tukey$Comparison<-factor(rownames(tukey),levels=rev(rownames(tukey)))
tukey$Comparison<-factor(tukey$Comparison,levels=tukey$Comparison[order(tukey$Estimate)])

p.tukey<-ggplot(tukey,aes(y=Comparison,x=Estimate))+
	geom_vline(xintercept=0,linetype = 2,col="red")+
	geom_errorbarh(aes(xmin=lwr,xmax=upr),height=.25,col="grey")+
	geom_point()+
	labs(x="Difference in proportions",title="95% family-wise confidence level")
p.tukey
```

The next figure shows the estimated proportions.

```{r}
lm2 <- lm(subscribed~treatmentF-1,data=bigdat)
source("confintHC.R") ## our own CI maker with HC SEs
theCIs<-confint(lm2,vcov=vcovHC(lm2,type="HC2"))
```


```{r}
theCIs<-as.data.frame(theCIs)
theCIs$Treatment<-c("Opt-in; List",
		      "Active; List",
		      "Opt-in; Quiz",
		      "Active; Quiz",
		      "Opt-in",
		      "Active"
		      )
theCIs$pbar <- coef(lm2)
theCIs$Treatment <- factor(theCIs$Treatment, levels=theCIs$Treatment[order(theCIs$pbar)])
names(theCIs)[1:2] <- c("ll","ul")
```

```{r}
# Plot with ggplot2
dodge<-position_dodge(width = 0.9)
p.props<-ggplot(theCIs,aes(Treatment,pbar,fill=Treatment))+
	geom_bar(stat = "identity", position = dodge)+
	geom_errorbar(aes(ymin=ll,ymax=ul),position = dodge,width=.25)+
	labs(list(y="Proportion Devices New on the Site",
		  title="Proportion New Devices by Treatment Group"))+
theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p.props
```



