

This file is mostly to record the series of steps we took to clean and collapse the Google Analytics data.

```{bash, eval=FALSE}
cp SBST-Batch-SendTracking-GAStats-15Aug2016.csv orig.csv
sed -i "" "s/\*no //g" SBST-Batch-SendTracking-GAStats-15Aug2016.csv
## By Hand: Added a header
sed -i "" 's/\/15,.*$/\/15,/' SBST-Batch-SendTracking-GAStats-15Aug2016.csv
sed -i "" 's/\/16,.*$/\/16,/' SBST-Batch-SendTracking-GAStats-15Aug2016.csv
gsed -i.bak '/^[0-9]/! s/^/,/' SBST-Batch-SendTracking-GAStats-15Aug2016.csv ## add extra column, use gnu-sed from brew install gnu-sed
gsed -i.bak 's/reported/\/ email/g' SBST-Batch-SendTracking-GAStats-15Aug2016.csv
```

The version of `SBST-Batch-SendTracking-GAStats-15Aug2016.csv` that is on Google Drive was cleaned as above.


```{r}
gadat<-read.csv("SBST-Batch-SendTracking-GAStats-15Aug2016.csv",as.is=TRUE,header=TRUE)
gadat$Source...Medium[gadat$Source...Medium==""]<-NA
table(gadat$Source...Medium)
```

```{r}
library(dplyr)
tabdat <- gadat %>% group_by(Source...Medium) %>% summarise(sum(New.Users,na.rm=TRUE))
names(tabdat) <- c("treatment","newusers")
tabdat <- tabdat[-7,]
wrkdat<-read.csv("data/wrkdat.csv",as.is=TRUE)
assigndat<-wrkdat %>% group_by(treatment) %>% summarize(sent=n())

dat <- cbind(assigndat,newusers=tabdat$newusers)
stopifnot(tabdat[tabdat$treatment=="newsletter_a / email","newusers"] == dat[dat$treatment=="A","newusers"])

dat$success <- dat$newusers
dat$failures <- dat$sent - dat$newusers

mat <- as.matrix(dat[,c("success","failures")])
rownames(mat) <- as.character(dat$treatment)

```

Expand these tables to make analyses easier with existing software:

```{r}
library(reshape2)
bigdat <- melt(apply(mat,1,function(x){ c(rep(1,x[1]),rep(0,x[2])) }))
names(bigdat) <- c("subscribed","treatment")
bigdat$subscribedF <- factor(bigdat$subscribed,levels=c(1,0))
bigdat$treatmentF <- factor(bigdat$treatment)
tab2<-with(bigdat,table(treatment,subscribedF))


```

Now, assess the hypotheses of no difference among treatments.

```{r}
prop.test(tab2)

```

Look at pairwise differences:

```{r}
library(multcomp)
thelm <- lm(subscribed~treatmentF,data=bigdat)
c(coef(thelm)[1], coef(thelm)[1]+coef(thelm)[2:6])
coeftest(thelm,vcov=vcovHC(thelm,type="HC2"))
allcomps <- glht(thelm,linfct=mcp(treatmentF="Tukey"))
summary(allcomps)
allcompsci <- confint(allcomps) ## adjusted for multiple comparisons
```

Now calculate CIs for the proportions:

```{r}

library(sandwich)
library(lmtest)
lm2 <- lm(subscribed~treatmentF-1,data=bigdat)
source("confintHC.R") ## our own CI maker with HC SEs
theCIs<-confint(lm2,vcov=vcovHC(lm2,type="HC2"))
```



