---
title: "DoD Military Community and Family Policy Military OneSource eNewsletter"
author: "Paul Testa"
date: "July 8, 2016"
output: 
    html_document:
        toc: TRUE
---

```{r init,echo=F}
library(knitr)
opts_chunk$set(eval=T,echo=F,results="hide",message=F,warning=F,cache=T)

```

\tableofcontents

# Setup

- Download "designdata.csv" and "subscriptiondata.csv" onto local machine
- Load files and libaries

```{r setup}
data<-read.csv("data/designdata.csv")
subscribers<-read.csv("data/subscriptiondata.csv")

```


# Data

```{r clean}
dim(data) # 491879 observations

# Make sure emails are unique:
length(unique(data$emailhash))==dim(data)[1]
length(unique(subscribers$emailhash))==dim(subscribers)[1]

dim(subscribers) # 7759 subscriptions
sum(data$emailhash%in%subscribers$emailhash) # 5563 subscribers in initial email list

# Rename Treatment indicator
names(data)[3]<-"treatment"

# Create indicator for multiple names
# Data
multi_names<-names(which(table(data$name1hash)>1))
data$multi_name01<-ifelse(data$name1hash%in%multi_names,1,0)
table(data$multi_name01) # 152828 have same names
# Subscribers
multi_names_sub<-names(which(table(subscribers$name1hash)>1))
subscribers$multi_name01<-ifelse(subscribers$name1hash%in%multi_names_sub,1,0)
table(subscribers$multi_name01) # 142 have same names

# Distribution of multiple names: Data
no_of_multi<-sort(unique(table(data$name1hash)))
counts<-hist(table(data$name1hash),plot=F,breaks=0:93)$count
counts<-counts[counts!=0]
tab_names_data<-cbind(c(no_of_multi,"No Name"),
                          c(counts,sum(is.na(data$name1hash))),
                          c(counts,sum(is.na(data$name1hash)))*c(no_of_multi,1)
                          )
colnames(tab_names_data)<-c("Times Name Appears","Frequency","Observations")
# Distribution of multiple names: Subscribers
no_of_multi_sub<-sort(unique(table(subscribers$name1hash)))
counts_sub<-hist(table(subscribers$name1hash),breaks=0:4,plot=F)$count
counts_sub<-counts_sub[counts_sub!=0]
tab_names_sub<-cbind(c(no_of_multi_sub,"No Name"),
                          c(counts_sub,sum(is.na(subscribers$name1hash))),
                          c(counts_sub,sum(is.na(subscribers$emailhash)))*c(no_of_multi_sub,1)
                          )
colnames(tab_names_sub)<-c("Times Name Appears","Frequency","Observations")


# Matches by email

sum(data$emailhash%in%subscribers$emailhash) # 5563

# Matches by name1hash
sum(data$name1hash%in%subscribers$name1hash) # 7951 subscribers in 




# Additional variables
summary(data$Order.Within.Condition)
table(data$Condition.Order.Within.Batch)


# DV Coding:
data$sub01_email<-ifelse(data$emailhash%in%subscribers$emailhash,1,0)
table(data$sub01_email)
sum(is.na(data$name1hash[data$sub01_email==1]))
# None of the matched emails have hashed names
data$sub01_name<-ifelse(data$name1hash%in%subscribers$name1hash,1,0)
table(data$sub01_name) # 7951 matches
# Nearly all of the hashed names are matched 

table(data$sub01_name,data$multi_name01) # 1700 are unique indivudals, 6251 are multipele cases



data$sub01_emailname<-ifelse(data$emailhash%in%subscribers$emailhash|
                                 data$name1hash%in%subscribers$name1hash[subscribers$multi_name01==0],1,0)

table(data$sub01_emailname) # Over matching...

# Throw names that occur more than once
data$sub01_emailname[data$multi_name01==1]<-0
table(data$sub01_emailname)
table(data$sub01_emailname,data$multi_name01)
table(data$sub01_emailname,data$sub01_email)

# Use this as the primary DV

data$subscribe01<-data$sub01_emailname


# Reorder as factor for prop tests
data$subscribe<-ifelse(data$subscribe01==1,"Subscribed","Not Subscribed")
data$subscribe<-factor(data$subscribe,levels=c("Subscribed","Not Subscribed"))
```


The primary dataset in this analysis contains `r dim(data)[1]` study participants, each randomly assigned to receive one of six treatment emails

- `r dim(data)[1]` participants
- `r dim(subscribers)[1]` subscribers
- `r sum(data$emailhash%in%subscribers$emailhash)` participants in list of subscribers
    - `r dim(subscribers)[1]-sum(data$emailhash%in%subscribers$emailhash)` emails in list of subcriber different from emails in list of participants


# Outcome


The outcome is whether participants subscribed to the newsletter. It was constructed in the following way: First subscribers were identified by matching their unique email address in the design dataset to a list of  `r dim(subscribers)[1]` subscribers' emails.  Additional subscribers were identified by then matching on first and last names in each dataset. Some names appear multiple times in each dataset. These cases, where a single subscriber name is matched to multiple participants with the same name, are then recoded as zeros^[Should they be recoded as NA? Maybe a bounds analysis, randomly assigning one person as the subscriber, but perhaps not necessary]. In total, using email address and names, we are able to match `r sum(data$subscribe01==1)` participants to the list of subscribers (i.e. `r round( sum(data$subscribe01==1)/dim(subscribers)[1]*100,1)`\% of subscribers).





# Randomization Assessment

```{r ra,results="asis"}
tab0<-data.frame(Treatement=names(table(data$treatment)),
                                  N=matrix(table(data$treatment)))
kable(tab0)
```

- Treatment appears to have been administred in 82 batches of 6000 participants with the exception of batch 82, which was sent  `r table(data$Batch)[82]` recipients. 


# Comparisons Across Newsletters

```{r pwcomp}
# Holm
pw_props_holm<-with(data,pairwise.prop.test(table(treatment,subscribe01)[,c(2,1)],
                                       p.adjust.method = "holm"
                                       ))
# No correction
pw_props_none<-with(data,pairwise.prop.test(table(treatment,subscribe01)[,c(2,1)],
                                       p.adjust.method = "none"
                                       ))
```


The tables below show:

- The raw counts and proportions of people subscribing by treatment status
- The p-values from pairwise comparisons of these proportions using the Holm (1979) correction for multiple comparisons, as well as wihtout any adjustments.

Overall subscription rates are low (~1 percent). The differences between conditions are also small (about one to two tenths of a percent). Using the Holm correction for multiple comparisons, `r sum(pw_props_holm$p.value<0.05,na.rm=T)` of the 15 comparisons are statistically significant (p<0.05, 8 p<0.06, and  `r sum(pw_props_none$p.value<0.05,na.rm=T)` without adjusting for multiple comparisons). 

## Proportion Subscribing  by Treatment Status
```{r tab1,results="asis"}
# Proportion Subscribing by Treatment
tab1<-rbind(table(data$treatment),
            table(data$subscribe,data$treatment),
            paste("**",round(unlist(lapply(split(data$subscribe01,f=data$treatment),mean)),4),"**",sep=""))
rownames(tab1)<-c("N","Subscribed","Not Subscribed","**Proportion Subscribing**")
kable(tab1,caption="Proprortion Subscribing by Treatment Status")
```

## P-Values for Pairwise Comparison of Proportions (Holm Correction)

```{r tab2,results="asis"}



tab_pw_props_holm<-matrix(sprintf("%.4f",round(pw_props_holm$p.value,5)),5,5)
colnames(tab_pw_props_holm)<-colnames(pw_props_holm$p.value)
rownames(tab_pw_props_holm)<-rownames(pw_props_holm$p.value)
tab_pw_props_holm[tab_pw_props_holm=="NA"] <- ""

kable(tab_pw_props_holm,
      caption="Pairwise comparisons using Pairwise comparison of proportions (Holm Correction)")

```


## P-Values for Pairwise Comparison of Proportions (No Correction for Multiple Comparisons)

```{r tab3,results="asis"}

tab_pw_props_none<-matrix(sprintf("%.4f",round(pw_props_none$p.value,5)),5,5)
colnames(tab_pw_props_none)<-colnames(pw_props_none$p.value)
rownames(tab_pw_props_none)<-rownames(pw_props_none$p.value)
tab_pw_props_none[tab_pw_props_none=="NA"] <- ""

kable(tab_pw_props_none,
      caption="Pairwise comparisons using Pairwise comparison of proportions (Holm Correction)")


```

# Comparisons Across Treatment Type


```{r tablemaker}
# Function to make tables 
table.maker<-function(x){
    # Difference of Proportions
     test<-stats::prop.test(table(x,data$subscribe01)[,c(2,1)])
     diff<- sprintf("%.4f",round(test$estimate[1]-test$estimate[2],5))
     pval<- sprintf("%.4f",round(test$p.value,5))
    # MH Test using Batch
    mhtest<-mantelhaen.test(table(x,data$subscribe,data$Batch))
    mhtest.pval<-sprintf("%.3f",round(mhtest$p.value,3))
    mhtest.stat<-sprintf("%.3f",round(mhtest$statistic,3))
    tab<-rbind(table(x),
            table(data$subscribe,x),
            paste("**",round(unlist(lapply(split(data$subscribe01,f=x),mean)),4),"**",sep=""),c(" "," "),
            c("*Statistic*","*p-value*"),
            c(diff,pval),
            c( mhtest.stat,mhtest.pval))
    rownames(tab)<-c("**Sample**","Subscribed","Not Subscribed","Proportion Subscribing","___","**Test**","Difference in Proportions"," CMH $\\chi^2$ Test")
    colnames(tab)<-paste("**",colnames(tab),"**",sep="")
    return(tab)
    
    
}
```

This section makes the following comparisons:

- A,C,E versus B,D,F (a test of opt-in versus enhanced active).
- A,B,C,D versus E,F (a test of Block of 10 present versus active)
- A,B versus C,D (a test of list versus quiz format)

For each comparison, the tables present:

- The raw counts and proportions of people subscribing by treatment status
- A test of the difference in proportions
- A test of the difference using the Cochran-Mantel-Haenszel Chi-Squared Test for Count Data treating each batch as separate strata.

Ovearll rates of subscription are higher:

- In treatment conditions with enhanced active choice (B,D,F) rather than a single opt-in (A,C,E)
- In treamtn conditions where the block 10 is absent (E,F) rather than present (A,B,C,D)
- In treament conditions that present information in a list format (A,B) rather than a quiz format (C,D).

# A,C,E versus B,D,F (a test of opt-in versus enhanced active).

```{r tabbdf, results="asis"}
# A,C,E versus B,D,F (a test of opt-in versus enhanced active).

data$treat_bdf_01<-as.numeric(grepl("B|D|F",data$treatment))
data$treat_bdf<-ifelse(grepl("B|D|F",data$treatment),"Enhanced Active","Opt-In")
tab_bdf<-table.maker(data$treat_bdf)
kable(tab_bdf,caption="Enhanced Active versus Single Opt-In")

```

# A,B,C,D versus E,F (a test of Block of 10 present versus active)

```{r tababcd,results="asis"}
# A,B,C,D versus E,F (a test of Block of 10 present versus active)

data$treat_abcd_01<-as.numeric(grepl("A|B|C|D",data$treatment))
data$treat_abcd<-ifelse(grepl("A|B|C|D",data$treatment),"Present","Absent")
data$treat_abcd<-factor(data$treat_abcd,levels=c("Present","Absent"))
tab_abcd<-table.maker(data$treat_abcd)
kable(tab_abcd,caption="Block of 10 Present or Absent")


```

# A,B versus C,D (a test of list versus quiz format)

```{r tabab,results="asis"}
# A,B versus C,D (a test of list versus quiz format)

data$treat_ab_01<-as.numeric(grepl("A|B",data$treatment))
data$treat_ab_01[!grepl("A|B|C|D",data$treatment)]<-NA
data$treat_ab<-ifelse(data$treat_ab_01==1,"List","Quiz")

tab_ab<-table.maker(data$treat_ab)
kable(tab_ab)

```

# Instances of Multiple Names

## Design Data

```{r, results="asis"}
kable(tab_names_data)
```

## Subscriber Data

```{r, results="asis"}
kable(tab_names_sub)
```


# Code

```{r, echo=T,eval=F}
<<init>>
<<setup>>
<<clean>>
<<ra>>
<<pwcomp>>
<<tab1>>
<<tab2>>
<<tab3>>
<<tablemaker>>
<<tabbdf>>
<<tababcd>>
<<tabab>>
```

